BFS search in miniKanren

### 2019-03-02 Paper/paper.tex
tuning

The \emph{cost} of an answer is the number of relation applications needed to find the answer. We use the miniKanren relation \texttt{repeato} in Fig.~\ref{def-repeato} to demonstrate the concept. We borrow this idea from Silvija Seres's work [*]. \texttt{repeato} relates a term \texttt{x} with a list whose elements are all \texttt{x}s. The \texttt{run} expression uses \texttt{repeato} to generate 4 lists whose elements are all \texttt{*}s. The order of the answers reflects the order miniKanren discovers them. The leftmost answer is discovered first. The order here is not suprising: to generate the answer \texttt{'()}, miniKanren needs to apply \texttt{repeato} only once. And it needs more applications of \texttt{repeato} to find the later answers that are more complicated. In this example, the cost of each answer is the same as one more than the number of \texttt{*}s. No answer of zero cost exists.

\begin{figure}
  \lstinputlisting{Figures/defrel-repeato.rkt}
  \lstinputlisting{Figures/run-repeato-1.rkt}
  \caption{\texttt{repeato}}
  \label{def-repeato}
\end{figure}

@@

\begin{figure}
	\lstinputlisting{Figures/run-conde-repeato-iDFS.rkt}
	\caption{run a program with interleaving depth-first search}
	\label{conde-repeato-iDFS}
\end{figure}

\begin{figure}
	\lstinputlisting{Figures/run-conde-repeato-BFS.rkt}
	\caption{run a program with breadth-first search}
	\label{conde-repeato-BFS}
\end{figure}

Although all answers by iDFS is not in cost-respecting order, the answers from each case are in cost-respecting order. The problem is that iDFS strategy prioritizes the first \texttt{conde} case considerablely. In general, when every conde case are equally productive, the iDFS strategy takes $1/2^{i}$ answers from the $i$-th case, except the last case, which share the same portion as the second last.

>>>>

The \emph{cost} of an answer is the number of relation applications needed to
find the answer. This idea is borrowed from Silvija Seres's work [*].
Now we illustrate the costs of answers by running a miniKanren relation.
Fig.~\ref{def-repeato} defines the relation \texttt{repeato} that 
relates a term \texttt{x} with a list whose elements are all \texttt{x}s.

\begin{figure}
  \lstinputlisting{Figures/defrel-repeato.rkt}
  \caption{\texttt{repeato}}
  \label{def-repeato}
\end{figure}

Consider the following \texttt{run} of \texttt{repeato}.
\begin{center}
  \begin{tabular}{c}
  \lstinputlisting{Figures/run-repeato-1.rkt}
   \end{tabular}
\end{center}

The above \texttt{run} generates 4 answers. All are lists of \texttt{*}s.
The order of the answers reflects the order miniKanren discovers them:
the leftmost answer is the first one. This result is not suprising:
to generate the first answer, \texttt{'()}, miniKanren needs to apply
\texttt{repeato} only once and the later answers need more recursive
applications. In this example, the cost of each answer is the same as
one more than the number of \texttt{*}s. The cost of \texttt{'()} is 1,
the cost of \texttt{'(*)} is 2, etc.

A list of answer is in the \emph{cost-respecting} order if no answer occurs
before another answer of a lower cost. In the above example, the answers are
cost-respecting. The iDFS search, however, does not generate cost-respecting
answers in general. As an example, consider the following \texttt{run} of
\texttt{repeato}.
\begin{center}
  \begin{tabular}{c}
   \lstinputlisting{Figures/run-conde-repeato-iDFS.rkt}
   \end{tabular}
\end{center}

The results are not cost-respecting. For example, \texttt{'(a a)} occurs before
\texttt{'(b)} while \texttt{'(a a)} is associated with a higher cost. 
The problem is that iDFS strategy prioritizes the first \texttt{conde} case
considerablely. In general, when every conde case are equally productive, the iDFS
strategy takes $1/2^{i}$ answers from the $i$-th case, except the last case,
which share the same portion as the second last.

@@

\lstinputlisting{Figures/run-conde-repeato-BFS.rkt}
